---
title: "p8105_hw3_yq2378"
author: "Qi Yumeng"
date: "2023-10-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn = -1)  # 禁用警告
```

```{r}
library(p8105.datasets)
library(tidyverse)
library(ggpubr)
```
# Problem 1
```{r}
data("instacart")
```
The data has `r nrow(instacart)` observations of `r ncol(instacart)` variables, where each row in the dataset is a product from an order. `order_id` is order identifier. `user_id` is customer identifier. `order_dow` is the day of the week on which the order was placed. `order_hour_of_day` is the hour of the day on which the order was placed. `product_name` is the name of the product. `aisle` is the name of the aisle. `department` is the name of the department. The following table gives a more detailed summary of the variables. 
```{r}
skimr::skim(instacart) 
```
There are 134 unique aisles and fresh vegetables is the most items ordered from. Fresh fruits comes the second and packaged vegetables fruit comes the third.
```{r}
instacart %>% group_by(aisle) %>% summarise(cnt = n()) %>% arrange(desc(cnt))
```
Grouped the data by aisle and reorder the data by the number of items ordered in each aisle, we could easily see that the number of fresh vegetables and fruits are almost twice as that of packaged vegetables fruits. The least popular items are dry pasta, oil vinegars and butter.

```{r}
instacart %>% group_by(aisle) %>% summarise(cnt = n()) %>%  filter(cnt>10000) %>% 
  ggplot(aes(x = reorder(aisle, cnt), y = cnt))+ geom_bar(stat="identity") + coord_flip() +
  geom_text(aes(label = cnt, y = 5000), cex = 2, col ='white')+
  xlab("Aisle") + theme_pubclean()
```

The three most popular items in “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
were listed below. For packaged vegetables, the hit products are Organic Baby Spinach, Organic Raspberries and Organic Blueberries. For baking ingredients, the hit products are Light Brown Sugar, 	Pure Baking Soda and Cane Sugar. For dog food care, the hit products are 	Snack Sticks Chicken & Rice Recipe Dog Treats, Organix Chicken & Brown Rice Recipe and Small Dog Biscuits
```{r}
instacart %>% filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits")) %>% group_by(aisle,product_name) %>% summarise(cnt = n()) %>% arrange(desc(cnt)) %>% slice(1:3)%>% arrange(desc(cnt))
```

The table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. For Coffee Ice Cream, the mean daily sales of Coffee Ice Cream is approximately 14.20 units. The highest daily sales occurred on Tuesday, with a value of 15.38 units. The lowest daily sales occurred on Friday, with a value of 12.26 units.As for Pink Lady Apples, the mean daily sales of Pink Lady Apples is approximately 12.54 units. The highest daily sales occurred on Wednesday, with a value of 14.25 units. The lowest daily sales occurred on Monday, with a value of 11.36 units

```{r}
instacart %>% filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>% group_by(order_dow,product_name) %>% summarise(mean_order_hour_of_day = round(mean(order_hour_of_day),2)) %>% pivot_wider(names_from = order_dow,values_from = mean_order_hour_of_day) %>% rename(Sun = "0", Mon = "1", Tue = "2", Wed = "3", Thu = "4",Fri = "5", Sat = "6")
```



# Problem 2


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
data("brfss_smart2010")
brfss_smart2010 = brfss_smart2010 %>% janitor::clean_names() %>% filter(topic == 'Overall Health' & response %in% c("Poor", "Excellent")) %>% 
  mutate(response = factor(response, levels = c("Poor","Excellent"), ordered = TRUE))
```

In 2002, 6 states, including PA, MA, NJ, CT, FL, NC were observed at 7 or more locations. In 2010, 14 states were left, including FL, NJ, TX, CA, MD, NC, NE,WA, MA,NY,OH, Co, PA, SC were observed at 7 or more locations. Details were listed below.
```{r}
brfss_smart2010 %>% filter(year == 2002) %>% group_by(locationabbr) %>% summarise( cnt = n_distinct(locationdesc)) %>% arrange(desc(cnt))  %>% filter(cnt >= 7)

brfss_smart2010 %>% filter(year == 2010) %>% group_by(locationabbr) %>% summarise( cnt = n_distinct(locationdesc)) %>% arrange(desc(cnt))  %>% filter(cnt >= 7)
```
The dataset `brfss_smart_excellent` contains 443 observations of 3 variables. From 2002 to 2010, mean data value across 51 states were presented by the "spaghetti" plot. Generally the lines were in a decreasing trend.

Distributions of `data_value` in different year and of different position is shown below.Clearly, for both 2006 and 2010, the Poor has lower data value compared to the Excellence. The max value of the poor never exceeds the Q1 of the Excellence. Rough little difference could be detetced between 2006 and 2010.

```{r}
brfss_smart_excellent = brfss_smart2010 %>% filter(response == "Excellent") %>% 
  group_by(year,locationabbr) %>% summarise(mean_data_value = mean(data_value))

brfss_smart_excellent %>% ggplot(aes(x = year, y = mean_data_value,group = locationabbr,col = locationabbr)) + 
  geom_line() + theme_pubr(legend = "right") + ylab("Mean Data Value") + xlab("Year")


brfss_smart2010 %>% filter(year %in% c(2006,2010)) %>%
  ggplot(aes(x = response, y = data_value))+ geom_boxplot()+ facet_grid(cols = vars(year)) + theme_pubr() +
  xlab("Response") + ylab("Data Value")

```

