---
title: "p8105_hw3_yq2378"
author: "Qi Yumeng"
date: "2023-10-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn = -1)  # 禁用警告
```

```{r}
library(p8105.datasets)
library(tidyverse)
library(ggpubr)
```
# Problem 1
```{r load p1 data}
data("instacart")
```
The data has `r nrow(instacart)` observations of `r ncol(instacart)` variables, where each row in the dataset is a product from an order. `order_id` is order identifier. `user_id` is customer identifier. `order_dow` is the day of the week on which the order was placed. `order_hour_of_day` is the hour of the day on which the order was placed. `product_name` is the name of the product. `aisle` is the name of the aisle. `department` is the name of the department. The following table gives a more detailed summary of the variables. 
```{r view data p1}
skimr::skim(instacart) 
```
There are 134 unique aisles and fresh vegetables is the most items ordered from. Fresh fruits comes the second and packaged vegetables fruit comes the third.
```{r aisles}
instacart %>% group_by(aisle) %>% summarise(cnt = n()) %>% arrange(desc(cnt))
```
Grouped the data by aisle and reorder the data by the number of items ordered in each aisle, we could easily see that the number of fresh vegetables and fruits are almost twice as that of packaged vegetables fruits. The least popular items are dry pasta, oil vinegars and butter.

```{r groupby and reorder}
instacart %>% group_by(aisle) %>% summarise(cnt = n()) %>%  filter(cnt>10000) %>% 
  ggplot(aes(x = reorder(aisle, cnt), y = cnt))+ geom_bar(stat="identity") + coord_flip() +
  geom_text(aes(label = cnt, y = 5000), cex = 2, col ='white')+
  xlab("Aisle") + theme_pubclean()
```

The three most popular items in “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
were listed below. For packaged vegetables, the hit products are Organic Baby Spinach, Organic Raspberries and Organic Blueberries. For baking ingredients, the hit products are Light Brown Sugar, 	Pure Baking Soda and Cane Sugar. For dog food care, the hit products are 	Snack Sticks Chicken & Rice Recipe Dog Treats, Organix Chicken & Brown Rice Recipe and Small Dog Biscuits
```{r most popular items}
instacart %>% filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits")) %>% group_by(aisle,product_name) %>% summarise(cnt = n()) %>% arrange(desc(cnt)) %>% slice(1:3)%>% arrange(desc(cnt))
```

The table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. For Coffee Ice Cream, the mean daily sales of Coffee Ice Cream is approximately 14.20 units. The highest daily sales occurred on Tuesday, with a value of 15.38 units. The lowest daily sales occurred on Friday, with a value of 12.26 units.As for Pink Lady Apples, the mean daily sales of Pink Lady Apples is approximately 12.54 units. The highest daily sales occurred on Wednesday, with a value of 14.25 units. The lowest daily sales occurred on Monday, with a value of 11.36 units

```{r pink&coffee}
instacart %>% filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>% group_by(order_dow,product_name) %>% summarise(mean_order_hour_of_day = round(mean(order_hour_of_day),2)) %>% pivot_wider(names_from = order_dow,values_from = mean_order_hour_of_day) %>% rename(Sun = "0", Mon = "1", Tue = "2", Wed = "3", Thu = "4",Fri = "5", Sat = "6")
```



# Problem 2


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r load data p2}
data("brfss_smart2010")
brfss_smart2010 = brfss_smart2010 %>% janitor::clean_names() %>% filter(topic == 'Overall Health' & response %in% c("Poor", "Excellent")) %>% 
  mutate(response = factor(response, levels = c("Poor","Excellent"), ordered = TRUE))
```

In 2002, 6 states, including PA, MA, NJ, CT, FL, NC were observed at 7 or more locations. In 2010, 14 states were left, including FL, NJ, TX, CA, MD, NC, NE,WA, MA,NY,OH, Co, PA, SC were observed at 7 or more locations. Details were listed below.
```{r states and location}
brfss_smart2010 %>% filter(year == 2002) %>% group_by(locationabbr) %>% summarise( cnt = n_distinct(locationdesc)) %>% arrange(desc(cnt))  %>% filter(cnt >= 7)

brfss_smart2010 %>% filter(year == 2010) %>% group_by(locationabbr) %>% summarise( cnt = n_distinct(locationdesc)) %>% arrange(desc(cnt))  %>% filter(cnt >= 7)
```
The dataset `brfss_smart_excellent` contains 443 observations of 3 variables. From 2002 to 2010, mean data value across 51 states were presented by the "spaghetti" plot. Generally the lines were in a decreasing trend.

Distributions of `data_value` in different year and of different position is shown below.Clearly, for both 2006 and 2010, the Poor has lower data value compared to the Excellence. The max value of the poor never exceeds the Q1 of the Excellence. Rough little difference could be detetced between 2006 and 2010.

```{r spaghettieplot}
brfss_smart_excellent = brfss_smart2010 %>% filter(response == "Excellent") %>% 
  group_by(year,locationabbr) %>% summarise(mean_data_value = mean(data_value))

brfss_smart_excellent %>% ggplot(aes(x = year, y = mean_data_value,group = locationabbr,col = locationabbr)) + 
  geom_line() + theme_pubr(legend = "right") + ylab("Mean Data Value") + xlab("Year")


brfss_smart2010 %>% filter(year %in% c(2006,2010)) %>%
  ggplot(aes(x = response, y = data_value))+ geom_boxplot()+ facet_grid(cols = vars(year)) + theme_pubr() +
  xlab("Response") + ylab("Data Value")

```
# Problem3
```{r load data p3}
# Load the data from the URL
demographic_url <- "https://p8105.com/data/nhanes_covar.csv"
accelerometer_url = "https://p8105.com/data/nhanes_accel.csv"
demographic = read.csv(demographic_url, skip = 4)
acceleromete = read.csv(accelerometer_url)
```
After loading, tidying, merging,filtering and organizing the data sets. The wide dataset, `df_merge` contains 228 observations of 1445 variables. The long dataset, `df_merge_long` contains 328329 observations with 7 variables.

```{r mergedata}
df_merge = demographic %>% filter(age >= 21) %>% na.omit() %>% 
  mutate(sex = factor(sex, levels = c(1,2), labels = c("male","female")),
         education = factor(education, levels = c(1,2,3), labels = c("Less than high school", "High school equivalent","More than high school"))) %>% left_join(acceleromete, by = "SEQN") %>% janitor::clean_names() 

df_merge_long = df_merge %>% pivot_longer(cols = min1:min1440,names_to = 'mims_tag',values_to = 'mims_value' )
```

The numbers of men and women with education levels categorized as "Less than high school" and "More than high school" are roughly equal, with 27 men and 28 women for the former and 56 men and 59 women for the latter. However, for "High school equivalent," there are more men than women, with 35 men and 23 women.

Regarding the distribution of age among different sex and education groups, men and women exhibit different patterns. For men, as age decreases, education level tends to rise. Both the median and the first and third quartiles (Q1 and Q3) decrease as education level increases. The range for each education level remains fairly consistent.

On the other hand, among women, those with "More than high school" education tend to be younger, while the median age for "High school equivalent" is slightly higher than that of "Less than high school." The range for "Less than high school" is narrower compared to the other two categories. Under the same education level, the medians of age are similar for "Less than high school" and "More than high school." However, for "High school equivalent," women clearly have a higher median age.

```{r boxplot}
df_merge %>% group_by(sex,education) %>% summarise(cnt = n())

df_merge %>% ggplot(aes(x = sex, y = age, fill = education)) + geom_boxplot() + theme_pubr(legend = "right")
```

We generated plots to analyze the relationship between total activities and age, making comparisons between males and females. In general, there is a negative correlation between age and total activities across all three education groups, albeit with some minor fluctuations. The trend lines further highlight this correlation, revealing that as education levels increase, the slope of the line becomes shallower for both males and females. Additionally, females exhibit a higher intercept and a greater absolute value of slope compared to males.

Remarkably, when examining the age range from 20 to 80, we observe that only within the 'less than high school' education group do the trend lines for males and females intersect. This intersection suggests a unique relationship in this particular group, different from the other education level

```{r smooth}
df_merge = df_merge %>% mutate(total_min = rowSums(select(., starts_with("min"))))

ggplot(df_merge, aes(x = age, y = total_min, col = sex)) + geom_line(alpha = 0.5) + facet_grid(cols = vars(education))+ theme_pubr(legend = "right") + ylab("total activities") + geom_smooth(method = "lm", se = FALSE, linetype = 2, linewidth = 0.3)
```

In a 24 hour time, generally speaking, as time increases the 24 hour activity time increases as well. For all three education level and both female and male, independent variable seem to be postively correlated to the dependent variables. As for the slope, there is little difference between gender. But for the More than High School group, the slop is more steep compared to the other two groups. As for the intercept, female has larger intercept than male only in the More than High School group
```{r geomline}

value = c(colSums(df_merge %>% 
                    filter(sex == "female" & education == "Less than high school") %>% 
                    select(starts_with("min"))),
          colSums(df_merge %>% 
                    filter(sex == "female" & education == "High school equivalent") %>% 
                    select(starts_with("min"))),
          colSums(df_merge %>% 
                    filter(sex == "female" & education == "More than high school") %>% 
                    select(starts_with("min"))),
          colSums(df_merge %>% 
                    filter(sex == "male" & education == "Less than high school") %>% 
                    select(starts_with("min"))),
          colSums(df_merge %>% 
                    filter(sex == "male" & education == "High school equivalent") %>% 
                    select(starts_with("min"))),
          colSums(df_merge %>% 
                    filter(sex == "male" & education == "More than high school") %>% 
                    select(starts_with("min"))))



df_merge_24 = data_frame( time = rep(seq(1,1440),times = 6),
                          sex = rep(c("female","male"), each = 1440*3),
                          education = rep(rep(c("Less than high school", 
                                                 "High school equivalent",
                                                 "More than high school"),each = 1440), times = 2),
                          value = value)


                       
ggplot(df_merge_24, aes(x = time, y = value, col = sex)) + geom_line(alpha = 0.5) + facet_grid(rows = vars(education))+ theme_pubr(legend = "right") + ylab("24-hour activity time") + geom_smooth(method = "lm", se = FALSE, linetype = 2, linewidth = 0.3) + ylab("24 Hour Activity Time") + xlab("Time")
```


